Data preparation

### 2023-03-16 04. Analysis/scripts/01_data_preparation.Rmd
Update 01_data_preparation.Rmd
Addressed nulls in position categories, supervisor corrections, and supervisor email addresses, added a check that they are addressed.

@@

@@

@@

@@

>>>>

Let's investigate whether these employees ever have other values for `r problem_variable %>% pull(variable)`.

@@

prep_01 %>% 
  DT::datatable(class = 'cell-border stripe')
```

They do have some.

Now let's look at how they change.

prep_02 %>% 
  DT::datatable(class = 'cell-border stripe')
```

It apprears that most start with nulls then became a category.

Let's check their position category at the earliest date and count.

```{r}
prep_02 %>% 
  group_by(person_id) %>% 
  filter(date_first == min(date_first)) %>% 
  ungroup() %>% 
  count(position_category_code,position_category_description)
```

Most of this group are null at the beginning. Some not.

Let's check the position category at the latest date and count.

```{r}
prep_02 %>% 
  group_by(person_id) %>% 
  filter(date_first == max(date_first)) %>% 
  ungroup() %>% 
  count(position_category_code,position_category_description)
```

Yes, but we want to use this in the analysis. We will fill the variable the most proximate value and create an indicator that these values were imputed.

```{r}
prep_xx <- 
  prep_building %>% 
  select(date, person_id,all_of(problem_variable %>% pull(variable))) %>% 
  distinct() %>% 
  mutate(
    position_category_imputed = case_when(is.na(position_category_code) ~ 1,
                                          is.na(position_category_description) ~ 1,
                                          TRUE ~ 0) 
  ) %>% 
  arrange(person_id, date) %>% 
  group_by(person_id) %>% 
  fill(position_category_code, .direction = "downup") %>% 
  fill(position_category_description, .direction = "downup") %>% 
  ungroup() %>% 
  mutate(
    across(
      .cols = all_of(problem_variable %>% pull(variable)),
      .fns = ~case_when(is.na(.) ~ "(Missing)",
                        TRUE ~ .)
    )
  )
  )

Compare against the identified cases above.

```{r}
prep_xx %>% 
  inner_join(prep_01,
            by =  c("person_id", "date"),
            suffix = c("_xx","_01")
  ) %>% 
  DT::datatable(class = 'cell-border stripe')
```

Looks good.

#### Step `r step_n`: `supervisor_correction`

```{r}
prep_building %>% 
  group_by(across(all_of(problem_variable %>% pull(variable)))) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
   flextable() %>% 
  theme_zebra() %>% 
  autofit()
```

Some increase in the use of this over time.

Let's count this along side `daily_status`.

```{r}
prep_building %>% 
  group_by(across(all_of(problem_variable %>% pull(variable))),daily_status) %>% 
  count() %>% 
   flextable() %>% 
  theme_zebra() %>% 
  autofit()
```

Yes, but we want to use this in the analysis. We will use with `daily_status` to define status.

```{r}
prep_xx <- 
  prep_building %>% 
  select(date, person_id,all_of(problem_variable %>% pull(variable)),daily_status) %>% 
  distinct() %>% 
  mutate(
    daily_status_imputed = case_when(is.na(supervisor_correction) &
                                       daily_status == "Supervisor Corrected" ~ "(Missing)",
                                     is.na(supervisor_correction) ~ daily_status,
                                     daily_status == "Supervisor Corrected" ~ supervisor_correction
                                     ) 
  )

Let's count these variables to make sure they are coming out correctly.

```{r}
prep_xx %>% 
  count(supervisor_correction,daily_status,daily_status_imputed) %>% 
  flextable() %>% 
  theme_zebra() %>% 
  autofit()
```

Looks good.

#### Step `r step_n`: `supervisor_e_mail_address`

Not too many.

Let's see if we can find these names in the roster data.

```{r}
nulls_01 %>% 
  select(supervisor_display_name) %>% 
  distinct() %>% 
  separate(supervisor_display_name,c("supervisor_first_name","supervisor_last_name"), sep = " ") %>% 
  mutate(
    full_name = glue::glue("{supervisor_last_name},{supervisor_first_name}")
  ) %>% 
  inner_join(prep_base_2)
```

Let's see if anyone has the same last name.

```{r}
prep_01 <- 
  nulls_01 %>% 
  select(supervisor_display_name) %>% 
  distinct() %>% 
  separate(supervisor_display_name,c("supervisor_first_name","supervisor_last_name"), sep = " ",
           remove = FALSE) %>% 
  mutate(
    full_name = glue::glue("{supervisor_last_name},{supervisor_first_name}")
  ) %>% 
  inner_join(prep_base_2 %>% 
               separate(full_name,c("supervisor_last_name","supervisor_first_name"), sep = ","),
             by = "supervisor_last_name") %>% 
  select(starts_with("supervisor_"), employee_e_mail_address,everything())

@@

This looks like the right person. Let's collect her information.

```{r}
prep_02 <- 
  prep_01 %>% 
  select(supervisor_display_name.x,employee_e_mail_address) %>% 
  distinct() %>% 
  filter(!is.na(employee_e_mail_address)) %>% 
  rename(supervisor_display_name = supervisor_display_name.x,
         supervisor_e_mail_address = employee_e_mail_address)

@@

```{r}
prep_xx <- 
  prep_building %>% 
  select(date, person_id,all_of(problem_variable %>% pull(variable)),supervisor_display_name) %>% 
  distinct() %>% 
  left_join(prep_02,by = "supervisor_display_name") %>% 
  mutate(
    supervisor_e_mail_address.z = 
      case_when(is.na(supervisor_e_mail_address.x) ~ supervisor_e_mail_address.y,
                TRUE ~supervisor_e_mail_address.x
                                     )
  )

Let's make sure it only effected this case.

```{r}
prep_xx %>% 
  filter(supervisor_e_mail_address.y == supervisor_e_mail_address.z) %>% 
  count(supervisor_display_name)
  arrange(desc(supervisor_e_mail_address.z))
```

Looks good. Let's only use `person_id`, `date`, and rename `supervisor_e_mail_address.z`.

```{r}
prep_xx <- 
  prep_xx %>% 
  select(person_id,date,supervisor_e_mail_address.z) %>% 
  rename(supervisor_e_mail_address = supervisor_e_mail_address.z)

@@

#### Step `r step_n`: Done!

Done.

### Combining  data

#### Create dataframe off non-null columns

Collect column names of variables not creating duplicates.

```{r}
non_problem_variables <- 
  column_meta_data  %>% 
  filter(!(blanks_by_variable | nulls_by_variable)) %>% 
  pull(variable)

non_problem_variables
```

Create dataframe of only these variables.

```{r}
prep_xx <- 
  prep_building %>% 
  select(any_of(non_problem_variables)) %>% 
  distinct()

```{r}
prep_building <- prep_problem %>% 
  left_join(prep_xx)
```

### Checking problems
